---
title: "PA1_template.Rmd"
author: "Meryem Sah"
date: "Sunday, December 14, 2014"
output: html_document
---


##1. Loading and preprocessing the data##

```{r, echo=TRUE}
data <- read.csv("activity.csv", header=TRUE, col.names = c("steps", "date", "interval"), colClasses=c("numeric","character","numeric"))

head(data)
summary(data)
str(data)

date <- as.Date(data$date, "%Y-%m-%d")
interval <- as.factor(data$interval)

```

###2. What is mean total number of steps taken per day?

```{r, echo=TRUE}
dailysteps <- aggregate(steps ~ date, data, FUN=sum, na.rm=TRUE)
hist(dailysteps$steps, breaks=10,ylab="Steps", xlab="Date", main="Steps per day", col="green")

mean(dailysteps$steps)
median(dailysteps$steps)
```

##3. What is the average daily activity pattern?

```{r, echo=TRUE}
dailyinterval <- aggregate(steps ~ interval, data, FUN=mean, na.rm=TRUE)
plot(dailyinterval$steps, type="l", col="blue", ylab="Steps", xlab="Interval")
```


###3.1 Maximum number of steps?

```{r, echo=TRUE}
dailyinterval$interval[which.max(dailyinterval$steps)]
```

##4. Imputing missing values

```{r, echo=TRUE}

sum(is.na(data))

datawithfilledNA <- merge(data, dailyinterval, by="interval", suffixes = c("",".new"))
filledNA <- is.na(datawithfilledNA$steps)
datawithfilledNA$steps[filledNA] <- datawithfilledNA$steps.new[filledNA]
datawithfilledNA <- datawithfilledNA[, c("steps", "date", "interval")]

head(datawithfilledNA)
sum(is.na(datawithfilledNA))

dailysteps <- aggregate(steps ~ date, datawithfilledNA, FUN=sum)
hist(dailysteps$steps, breaks=10,ylab="Steps", xlab="Date", main="Steps per day (filled NA)", col="green")

mean(as.integer(dailysteps$steps))
median(as.integer(dailysteps$steps))
  
```
The impact of the missing data is minimal. The mean remained the same and the histogram looks similar to the one without the missing data.The median changed and became a bit greater than the old one.

##5. Are there differences in activity patterns between weekdays and weekends?


```{r echo=TRUE}
daytype <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
daytype <- as.factor(sapply(datawithfilledNA$date, daytype))

```
